Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-05-01 tex/scam/reviews-v1/response-v1.tex

If the true regression function is a convex additive function, then additive faithfulness holds trivially--the best additive approximation to an additive function is the function itself. The estimation procedure becomes simpler--the DC stage is no longer necessary and can be safely removed.

In the finite sample analysis, the assumptions would not change except that A3 would be $\| f^*_k \|_\infty \leq B$ for all $k$ since $f_0 = f^*$ in this case. The signal strength definition can be made simpler: $\alpha^+ = \min_k \mathbb{E}( f^*_k(X_k)^2 )$. The rates would not improve.

\item \rc{If we actually wish to fit smooth convex univariate components, is
it easy to add this condition as a regularization term and incorporate
it into the analysis?}

Smoothness can be incorporated by using B-spline basis with uniformly spaced knots. The details are described in a 2014 paper by Pya and Wood which we have cited in the introduction. We chose not to discuss estimating smooth convex functions because it would introduce an additional tuning parameter.

The analysis may require much more work if one uses a B-spline basis in the estimation. B-splines may significantly complicate the false positive analysis, which rely on the particular form of the optimization. The false negative analysis may not be affected much because it mainly uses bracketing entropy.

@@

We have changed all ``F'' to ``P'' to represent the distribution, and, to avoid confusing the density with the dimensionality, we now write $p(x)$ instead of just $p$ whenever we refer to the density. We believe that this change, along with the context, are enough to disambiguate the two quantities.

\item \rc{I would suggest including the dependence on the positive-valued lower
bound on $p(x)$ (as assumed in (A5)) in the statement of Theorem 3. This
is a point worth commenting upon, since the earlier theorems only
require $p(x)$ to be strictly positive. Assumption (A5) seems like it
could also be interpreted as a sort of minimum signal strength
requirement for successful variable selection.}

\item \rc{It would be good to point to Assumption (A3) at the beginning of
Section 5 when discussing the $B$-boundedness condition. In the third
paragraph, the authors might note that the $B$-boundedness condition
arises from the analysis of Theorem 5.2. To clarify, the penultimate
paragraph on page 22 should state that Theorem 5.1 controls false
positives alone, whereas Theorem 5.3 (to follow) provides
finite-sample results controlling the false negative rate.}

We have made the suggested modifications except for the one in the second sentence. The $B$-boundedness condition is used in both Theorem 5.2 and Theorem 5.3. It is relied upon critically in Theorem 5.3 and used for convenience in Theorem 5.2.

@@

\item \rc{At the top
of page 23, the remark regarding mutual incoherence conditions is a
bit confusing. Aren’t the conditions imposed in Theorem 5.1 on the
covariates essentially analogous to the mutual incoherence conditions
in this case? }

The conditions in Theorem 5.1 are indeed analogous and we have edited the remark accordingly. The conditions in Theorem 5.1 are much more strict than the mutual incoherence condition though.

### 2015-05-01 tex/scam/additive_faithful.tex
boundary flat

We now show that under a general class of distributions which we
characterize below, convex multivariate functions are additively
faithful.

\begin{definition}
\label{defn:boundary-point}
Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$. $p(\mathbf{x})$ satisfies
the \emph{boundary flatness condition} if $0 < \inf_{\mathbf{x}} p(\mathbf{x}) \leq \sup_{\mathbf{x}} p(\mathbf{x}) < \infty$, and for all $j$, there exists some $\epsilon > 0$ such that for all $\mathbf{x}_{-j}$, and for all $x_j \in [0, \epsilon) \cup (1-\epsilon, 1]$, $p(\mathbf{x}_{-j} \given x_j)$ is twice differentiable in $x_j$, that $p(\mathbf{x}_{-j} \given x_j), \frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}, \frac{\partial^2 p(\mathbf{x}_{-j} \given x_j) }{\partial^2 x_j}$ are bounded, and that
\begin{equation}
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at } x_j = 0, x_j = 1
\end{equation}

@@

The boundary flatness condition intuitively states that
the conditional effect of $X_k$ weakens as $\mathbf{x}$ approaches the boundary. It is a weak condition because it affects only an $\epsilon$-small region around the boundary; $p(\mathbf{x}_{-j} \given x_j)$ can take arbitrary shapes away from the boundary; it can induce arbitrary correlation structure (provided $p(\mathbf{x}) >0$) and need not be continuous. Before presenting the main theorem, let us see some examples of boundary flat distributions.

\begin{example} The following densities satisfy boundary flatness.
\begin{enumerate}
\item If $\mathbf{p}$ is a product density, then boundary flatness is trivially
satisfied.
\item Let $p(\mathbf{x})$ be a joint density such that $0 < \inf_{\mathbf{x}} p(\mathbf{x}) \leq \sup_{\mathbf{x}} p(\mathbf{x}) < \infty$. Suppose $p(\mathbf{x})$ has a bounded second derivative for all $\mathbf{x}$. If, for all $j$,
$\partial_{x_j} p(x_j,\mathbf{x}_{-j}) =
\partial_{x_j}^2 p(x_j, \mathbf{x}_{-j}) = 0$ at $x_j = 0, 1$. Then it is possible to show that the two derivatives of $p(\mathbf{x}_{-j} \given x_j) = \frac{p(x_j, \mathbf{x}_{-j})}{p(x_j)}$ are 0 at $x_j = 0,1$; boundary flatness thus holds.
\item Suppose $p(\mathbf{x})$ is a bounded density over $[\epsilon, 1-\epsilon]^p$ for some $0 < \epsilon < 1/2$. Let $q(\mathbf{x})$ be a boundary flat density over $[0,1]^p$, take the uniform distribution for instance. Define a mixture $p_{\lambda, \epsilon}(\mathbf{x}) = \lambda q(\mathbf{x}) + (1-\lambda) p(\mathbf{x})$ where $0 < \lambda \leq 1$, then $p_{\lambda, \epsilon}(\mathbf{x})$ is boundary flat over $[0,1]^p$. Furthermore, $\epsilon$ and $\lambda$ can be made arbitrarily small and therefore, any bounded density over a hypercube can be approximated arbitrarily well by a boundary flat distribution.

% $$
% p'(\mathbf{x}_{-j} \given x_j) = \frac{ p(x_j, \mathbf{x}_{-j}) p'(x_j) - 
%    p'(x_j, \mathbf{x}_{-j}) p(x_j)}{ p(x_j)^2 } = 0 
% $$
% \begin{align*}
% p''(\mathbf{x}_{-j} \given x_j) = &
%     \frac{ p(x_j)^2 \left( p(x_j, x_{-j}) p''(x_j) - p''(x_j, x_{-j}) p(x_j) \right) - 2p(x_j)p'(x_j) \left(p(x_j, x_{-j})p'(x_j) - p'(x_j, x_{-j}) p(x_j)  \right) }{p(x_j)^4}
%  \end{align*}
\end{enumerate}
\end{example}

@@

\begin{theorem}
\label{thm:convex_faithful}
Let $p(\mathbf{x})$ be a positive density supported on $C=[0,1]^p$ that satisfies
the boundary flatness property. 
If $f$ is convex and continuously twice differentiable, then $f$ is additively faithful under $p(\mathbf{x})$.
\end{theorem}

@@

\begin{remark}
It is possible to get a similar result for distributions with
unbounded support, by using a limit condition $\lim_{|x_k| \rightarrow
  \infty} \frac{\partial p(\mathbf{\scriptstyle x}_{-k} \given x_k)}{\partial x_k}
= 0$.  Such a limit condition however is not obeyed by many common distributions such as the multivariate Gaussian distribution. The next example shows that certain convex functions are not additively faithful under certain multivariate Gaussian distributions.
\end{remark}

@@

Although the Gaussian distribution does not satisfy the boundary flatness condition, it is possible to approximate the Gaussian distribution arbitrarily well with distributions that do satisfy the boundary flatness conditions.

@@

\begin{theorem}
\label{thm:acdc_faithful}
Suppose $p(\mathbf{x})$ is a positive density on $C=[0,1]^p$ that
satisfies the boundary flatness condition. Suppose that $f$ is convex
and continuously twice differentiable. Let $f^*_k$ and $g^*_k$ be as defined in equations
\eqref{eqn:convex_additive} and \eqref{eqn:concave_postprocess}, then the $f^*_k$'s and the $g^*_k$'s are unique. Furthermore, 
$f^*_k = 0$ and $g^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x})
= 0$, that is, $f$ does not depend on~$x_k$. 
\end{theorem}

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(\mathbf{x})$ is a positive density on $C=[0,1]^p$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex and continuously twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be a bounded function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}

>>>>

We now show that under a general class of distributions which we
characterize below, convex multivariate functions are additively
faithful. To simplify presentation, we restrict our attention to densities bounded away from $0/\infty$, that is, $p(\mathbf{x})$ such that $0 < \inf p(\mathbf{x}) \leq \sup p(\mathbf{x}) < \infty$.

\begin{definition}
\label{defn:boundary-point}
Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$. We say that $p(\mathbf{x})$ satisfies the \emph{boundary flatness condition} if for all $j$, there exists some $\epsilon > 0$ such that for all $\mathbf{x}_{-j}$, and for all $x_j \in [0, \epsilon) \cup (1-\epsilon, 1]$, $p(\mathbf{x}_{-j} \given x_j)$ is twice differentiable in $x_j$, that $p(\mathbf{x}_{-j} \given x_j), \frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}, \frac{\partial^2 p(\mathbf{x}_{-j} \given x_j) }{\partial^2 x_j}$ are bounded, and that
\begin{equation}
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at } x_j = 0, x_j = 1
\end{equation}

@@

The boundary flatness condition intuitively states that two conditional densities $p(\mathbf{x}_{-j} \given x_j)$ and $p(\mathbf{x}_{-j} \given x_j')$ are similar when $x_j$ and $x_j'$ are both close to $0$ or both close to $1$. Product densities satisfy boundary flatness trivially. Boundary flatness is a weak condition because it affects only an $\epsilon$-small region around the boundary; $p(\mathbf{x}_{-j} \given x_j)$ can take arbitrary shapes away from the boundary. Boundary flatness also allows arbitrary correlation structure between the variables (provided $p(\mathbf{x}) >0$). In Section~\ref{sec:boundary_flat}, we give a detailed discussion of the boundary flatness condition and show many examples of boundary flat distributions; in particular, we show that any distribution supported on a compact set can be approximated arbitrarily well by boundary flat distributions.

@@

\begin{theorem}
\label{thm:convex_faithful}
Let $p(\mathbf{x})$ be a density supported on $C=[0,1]^p$ and bounded away from $0/\infty$ that satisfies
the boundary flatness property. 
If $f$ is convex and continuously twice differentiable, then $f$ is additively faithful under $p(\mathbf{x})$.
\end{theorem}

@@

\subsection{Boundary Flatness Examples}
\label{sec:boundary_flat}

In this section, we give more examples of boundary flat densities (see Definition~\ref{defn:boundary-point}) and discuss extending the notion of boundary flatness to densities with a more general support. We first start with an sufficient condition on the \emph{joint density} that ensures boundary flatness.

\begin{example}
\label{ex:joint_density_flat}
Boundary flatness is satisfied if the joint density becomes flat at the boundary. To be precise, let $p(\mathbf{x})$ be a joint density bounded away from $0/\infty$. Suppose, for all $j$,  $\partial_{x_j} p(x_j, \mathbf{x}_{-j})$ and $\partial^2_{x_j} p(x_j, \mathbf{x}_{-j})$ exist and are bounded for all $\mathbf{x}_{-j}$ and for all $x_j \in [0, \epsilon) \cup (1-\epsilon, 1]$ for some $\epsilon >0$.

Suppose also, for all $j$,
$$
\partial_{x_j} p(x_j,\mathbf{x}_{-j}) =
\partial_{x_j}^2 p(x_j, \mathbf{x}_{-j}) = 0\quad  \trm{at $x_j = 0, 1$,}
$$
then it is straightforward to show that $\partial_{x_j} p(\mathbf{x}_{-j} \given x_j) = \partial_{x_j}^2 p(\mathbf{x}_{-j} \given x_j) = 0$ at $x_j = 0,1$ as well. One need only see that the derivatives of the marginal densities are zero at the boundary and apply the quotient rule.
\end{example}

The next example shows that any bounded density over a hypercube can be approximated arbitrarily well by some boundary flat distributions.

\begin{example}
\label{ex:mixture_approx}
Suppose $p(\mathbf{x})$ is a bounded density over $[\epsilon, 1-\epsilon]^p$ for some $0 < \epsilon < 1/2$. Let $q(\mathbf{x})$ be a boundary flat density over $[0,1]^p$. Define a mixture $p_{\lambda, \epsilon}(\mathbf{x}) = \lambda q(\mathbf{x}) + (1-\lambda) p(\mathbf{x})$ where $0 < \lambda \leq 1$, then $p_{\lambda, \epsilon}(\mathbf{x})$ is boundary flat over $[0,1]^p$. Furthermore, $\epsilon$ and $\lambda$ can be made arbitrarily small and therefore, $p(\mathbf{x})$ can be approximated arbitrarily well (in $L_1$ for example) by a boundary flat distribution. 
\end{example}

The result in Example~\ref{ex:joint_density_flat} further shows that boundary flat densities can arise from monotone marginal transformations of more general random vectors. This gives another way of approximating general densities by boundary flat densities.

\begin{example}
Let $X = (X_1,...,X_p)$ be drawn from a density supported on $[0,1]^p$, bounded away from $0/\infty$, and with a bounded second derivative. Let $\phi : [0,1] \rightarrow [0,1]$ be a monotonically increasing transformation. We will show that under certain conditions on $\phi$, the density of $(\phi(X_1), ..., \phi(X_p))$ is boundary flat.

Suppose $\phi^{-1}$ has a bounded third derivative on some neighborhood around the boundary, say $[0, \epsilon) \cup (1-\epsilon, 1]$ and suppose $\phi^{-1\prime}(z) = \phi^{-1\prime\prime}(z) = \phi^{-1\prime\prime\prime}(z) = 0$ at $z=0,1$.

Then, the density $p(z_1,...,z_p)$ of the transformed variables $z_j = \phi(x_j)$ is 
$$ 
p(z_1,...,z_p) = p_X(\phi^{-1}(z_1), ..., \phi^{-1}(z_p)) \prod_{j=1}^p \phi^{-1\prime}(z_j),
$$ where $p_X(\cdot,...,\cdot)$ is the density of $X$.
One can use the product rule and the chain rule to differentiate $p(z_1,...,z_p)$ and verify that $\partial_{z_j} p(\mathbf{z}) = \partial^2_{z_j} p(\mathbf{z}) = 0$ at $z_j = 0,1$ and thus show that $p(z_1,...,z_p)$ satisfies boundary flatness by Example~\ref{ex:joint_density_flat}.

As a concrete example, we can use the following transformation $\phi_\epsilon$, defined in term of its inverse:
$$\phi_\epsilon^{-1}(z) = \left\{ \begin{array}{cc} z^4 & \trm{if $z \leq \epsilon$} \\ 
                     \frac{1-2\epsilon^4}{1-2\epsilon} z - \frac{ \epsilon + \epsilon^4}{1-2\epsilon} & \trm{if $\epsilon \leq z \leq 1-\epsilon$} \\
                      1-(1-z)^4 & \trm{if $z \geq 1-\epsilon$} \end{array} \right.$$
for any $1/2 > \epsilon > 0$. As $\epsilon$ becomes small, $\phi_\epsilon$ becomes close to the identity and so $(\phi_\epsilon(X_1), ..., \phi_\epsilon(X_p))$ closely approximates the original random variable $(X_1,...,X_p)$. Intuitively, $\phi$ stretches $X$ near the boundary so that changes in the transformed variable $z$ near the boundary corresponds to much smaller change in original variable $x$. 
\end{example}

In our discussion so far, we have restricted ourselves to densities supported and positive on the hypercube $[0,1]^p$ to minimize extraneous technical details. It may be possible to extend the analysis to densities whose support is a convex and compact set so long as the marginal density $p(x_j) > 0$ for all $x_j$ in the support.

It may also possible to extend similar result to densities with an
unbounded support, by using a limit condition $\lim_{|x_k| \rightarrow
  \infty} \frac{\partial p(\mathbf{\scriptstyle x}_{-k} \given x_k)}{\partial x_k}
= 0$.  Such a limit condition however is not obeyed by a correlated multivariate Gaussian distribution. 
The next example shows that certain convex functions are not additively faithful under certain multivariate Gaussian distributions.

@@

Although the Gaussian distribution does not satisfy the boundary flatness condition, it is possible to approximate the Gaussian distribution arbitrarily well with distributions that do satisfy the boundary flatness conditions. We use the similar idea as that of Example~\ref{ex:mixture_approx}.

@@

\begin{theorem}
\label{thm:acdc_faithful}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from $0/\infty$ that
satisfies the boundary flatness condition. Suppose that $f$ is convex
and continuously twice differentiable. Let $f^*_k$ and $g^*_k$ be as defined in equations
\eqref{eqn:convex_additive} and \eqref{eqn:concave_postprocess}, then the $f^*_k$'s and the $g^*_k$'s are unique. Furthermore, 
$f^*_k = 0$ and $g^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x})
= 0$, that is, $f$ does not depend on~$x_k$. 
\end{theorem}

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from $0/\infty$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex and continuously twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be a bounded function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}

### 2015-05-01 tex/scam/experiments.tex
boundary flat

The purpose of the simulations is to see how the probability of variable selection scales with the number of samples $n$, the ambient dimension $p$, and the sparsity level $s$. We study both the independent case where $P = N(0, I_p)$ and the correlated case where $P$ is a correlated Gaussian Copula modified slightly to satisfy boundary flatness. We measure the probability of exact selection in the independent case and the probability of screening in the correlated case. We also study both cases where the regression function is parametric (quadratic) and cases where the regression function is nonparametric (softmax of linear forms). In all our experiments, we mark a variable as selected if either the AC estimate $\| \hat{f}_j \|_\infty$ or the DC estimate $\| \hat{g}_k \|_\infty$ is larger than $10^{-6}$. We set $\lambda = 0.5 \sqrt{\frac{\log^2 np}{n}}$ for all the simulations.

@@

In the third and fourth simulation, we use a correlated design. We generate $X$ from a non-Gaussian boundary flat distribution with covariance $\Sigma$. The distribution we used is a mixture of a uniform distribution and a Gaussian Copula.
\[
X \sim \gamma U([-2, 2]^p) + (1-\gamma) \trm{Copula}(0, \Sigma, F)
\]
The Gaussian Copula is a way to customize the marginal distributions of a Gaussian random variable while maintaining the same covariance. Gaussian Copula results when one applies a monotone transformation $F^{-1} \Phi$ onto each of the variables of a Gaussian random vector where $\Phi$ is the normal CDF and $F$ is the CDF of the new marginal distribution. In all our experiments, we set $\gamma = 0.05$ and set the marginal CDF $F$ so that marginal density of the Copula is bimodal and supported on $[-1.8, 1.8]$. The resulting marginal density of the mixture is shown in Figure~\ref{fig:copula_marginal}. Notice that boundary flatness holds because the distribution is uniform in the boundary area $[-2,2]^p \backslash [-1.8, 1.8]^p$.

@@

In the fourth, fifth, and sixth simulation, we use a softmax function as the ground truth
\[
f_0(x_{iS}) = \log \left( \sum_{k=1}^K \exp( \beta_k^\tran x_{iS} ) \right) - \mu
\]
We generate random unit vectors as $\{\beta_k \in \R^s\}_{k=1,...,K}$ and choose $\mu$ so that $f_0$ has mean-zero. We set $K = 7$ for all the experiments.

@@

For the \textbf{fifth experiment} (Figure~\ref{Support}h), we vary the true sparsity level $s$. We let $f_0$ be the softmax function, $X \sim N(0, I)$, and set $p=128$. We compute the probability of exact recovery over 40 independent trials and plot the results in Figure~\ref{Support}(h).

For the \textbf{sixth experiment} (Figure~\ref{fig:ac_v_dc}), we compare the variables selected via the AC stage and the variables selected via the DC stage. We use the softmax regression function and $X$ drawn from the boundary flat mixture distribution with a Toeplitz covariance and correlation level $\nu = -0.7$. We set $s=5, n=500, p=128$. We perform 30 independent trials and plot the frequency of variable selection in Figure~\ref{fig:ac_v_dc}. The true variables are $X_j$ for $j=5,6,7,8,9,10$. We plot the frequency of selection among only the first 20 variables, that is, $X_j$ for $j=1,...,20$. We do not plot selection frequencies for variables 21 to 128 because they are almost never selected by either AC or DC. As can be seen, the DC stage is slightly helpful in recovering the true variables but its effect is not significant. We thus believe that the DC stage, though important in theory, is not as important in practice; it may be omitted without significant detriment to the overall result.

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cc}
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveT} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveR} \\
(a) quadratic $f_0$, independent $X$, varying $p$  & (b) quadratic $f_0$, independent $X$, varying $\Q$ \\
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/Q} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveC}  \\
(c) four $\Q$ matrices used in (b) & (d) quadratic $f_0$, correlated $X$, varying correlation $\nu$ \\
\includegraphics[width=.42\textwidth]{figs/C_support_box} & 
\includegraphics[width=.42\textwidth]{figs/CurveS}\\
(e) recovered support size for $\nu=0.5$ in (d) & (f) softmax $f_0$, correlated $X$, varying $p$ \\
\includegraphics[width=0.42\textwidth]{figs/S_support_box} & \\
(g) recovered support size for $p=256$ in (f) & (h) softmax $f_0$, correlated $X$, varying $s$ 
\end{tabular}
\end{center}
\caption{Support recovery results where the additive assumption is
  correct (a), incorrect (b), (c), and with correlated design (d).}\label{Support}
\vskip10pt
\end{figure*}

>>>>

We study both the independent case where $P = N(0, I_p)$ and the correlated case where $P$ is a correlated Gaussian Copula modified slightly to satisfy the boundary flatness condition. We measure the probability of exact selection in the independent case and the probability of screening in the correlated case. We also study both cases where the regression function is parametric (quadratic) and cases where the regression function is nonparametric (softmax of linear forms). In all our experiments, we mark a variable as selected if either the AC estimate $\| \hat{f}_j \|_\infty$ or the DC estimate $\| \hat{g}_k \|_\infty$ is larger than $10^{-6}$. We set $\lambda = 0.5 \sqrt{\frac{\log^2 np}{n}}$ for all the simulations.

@@

In the third, fourth, and fifth simulation, we use a correlated design. We generate $X$ from a non-Gaussian boundary flat distribution with covariance $\Sigma$. The distribution we used is a mixture of a uniform distribution and a Gaussian Copula.
\[
X \sim \gamma U([-2, 2]^p) + (1-\gamma) \trm{Copula}(0, \Sigma, F)
\]
The Gaussian Copula is a way to customize the marginal distributions of a Gaussian random variable while maintaining the same covariance. Gaussian Copula results when one applies a monotone transformation $F^{-1} \Phi$ onto each of the variables of a Gaussian random vector where $\Phi$ is the normal CDF and $F$ is the CDF of the new marginal distribution. In all our experiments, we set $\gamma = 0.05$ and set the marginal CDF $F$ so that marginal density of the Copula is bimodal and supported on $[-1.8, 1.8]$. The resulting marginal density of the mixture is shown in Figure~\ref{fig:copula_marginal}. Notice that boundary flatness holds because the distribution is uniform in the boundary area $[-2,2]^p \backslash [-1.8, 1.8]^p$.

@@

In the fourth and fifth simulation, we use a softmax function as the ground truth
\begin{align}
f_0(x_{iS}) = \log \left( \sum_{k=1}^K \exp( \beta_k^\tran x_{iS} ) \right) - \mu
\end{align}
We generate random unit vectors as $\{\beta_k \in \R^s\}_{k=1,...,K}$ and choose $\mu$ so that $f_0$ has mean-zero. We set $K = 7$ for all the experiments.

@@

%For the \textbf{fifth experiment} (Figure~\ref{Support}h), we vary the true sparsity level $s$. We let $f_0$ be the softmax function, $X \sim N(0, I)$, and set $p=128$. We compute the probability of exact recovery over 40 independent trials and plot the results in Figure~\ref{Support}(h).

For the \textbf{fifth simulation} (Figure~\ref{fig:ac_v_dc}), we compare the variables selected via the AC stage and the variables selected via the DC stage. We use the softmax regression function and $X$ drawn from the boundary flat mixture distribution with a Toeplitz covariance and correlation level $\nu = -0.7$. We set $s=5, n=500, p=128$. We perform 30 independent trials and plot the frequency of variable selection in Figure~\ref{fig:ac_v_dc}. The true variables are $X_j$ for $j=5,6,7,8,9,10$. We plot the frequency of selection among only the first 20 variables, that is, $X_j$ for $j=1,...,20$. We do not plot selection frequencies for variables 21 to 128 because they are almost never selected by either AC or DC. As can be seen, the DC stage is slightly helpful in recovering the true variables but its effect is not significant. We thus believe that the DC stage, though important in theory, is not as important in practice; it may be omitted without significant detriment to the overall result.

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cc}
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveT} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveR} \\
(a) quadratic $f_0$, independent $X$, varying $p$  & (b) quadratic $f_0$, independent $X$, varying $\Q$ \\
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/Q} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveC}  \\
(c) four $\Q$ matrices used in (b) & (d) quadratic $f_0$, correlated $X$, varying correlation $\nu$ \\
\includegraphics[width=.42\textwidth]{figs/C_support_box} & 
\includegraphics[width=.42\textwidth]{figs/CurveS}\\
(e) recovered support size for $\nu=0.5$ in (d) & (f) softmax $f_0$, correlated $X$, varying $p$ \\
%\includegraphics[width=0.42\textwidth]{figs/S_support_box} & \\
%(g) recovered support size for $p=256$ in (f) & (h) softmax $f_0$, correlated $X$, varying $s$ 
\end{tabular}
\end{center}
\caption{Support recovery results.}
\label{Support}
\vskip10pt
\end{figure*}

### 2015-05-01 tex/scam/sparsistency.tex
boundary flat

In our analysis, we assume that an upper bound $B$ to $\| \hat{f}_k \|_\infty$ 
is imposed in the optimization procedure, where $B$ is chosen to also upper bound $\| f^*_k \|_\infty$. This $B$-boundedness constraint is added so that we may use the convex function bracketing results from~\cite{kim2014global} to establish uniform convergence between the empirical risk and the population risk. We emphasize that this constraint is not needed in practice and we do not use it for any of our simulations.

@@

Theorem~\ref{thm:deterministic} allows us to separately analyze the false negative
rates and false positive rates. To control false positives,
we analyze the condition on $\lambda_n$ for $k \in S^c$. To control
false negatives, we need only analyze the restricted regression with only $|S|$ variables.

The proof of Theorem~\ref{thm:deterministic} analyses the KKT
conditions of optimization~\eqref{opt:alternate_opt}.  This parallels
the now standard \emph{primal-dual witness}
technique~\citep{wainwright2009sharp}. However, we cannot derive analogous
\emph{mutual incoherence} conditions because the estimation is
nonparametric---even the low dimensional restricted regression has
$s(n-1)$ variables. The details of the proof are given in
Section~\ref{sec:deterministic_proof} of the Appendix.

\subsection{Probabilistic Setting}

@@

\begin{theorem}[Controlling false positives]
\label{thm:false_positive}
Suppose assumptions A1-A5 hold. Define $\tilde{\sigma} \equiv \max(\sigma, B)$ and define $\mathsf{range}_k = X_{k \pi_k(n)} - X_{k \pi_k(1)}$. Suppose $p \leq O\big( \exp(c n) \big)$ and $n \geq C$ for some positive constants $0 < c <1$ and $C$.  Suppose also
\begin{equation}
\lambda_n \geq 512 s \tilde{\sigma}  \sqrt{ \frac{\log^2 np}{n}}.
\end{equation}  
Then with probability at least $ 1 - \frac{24}{n}$, for all $k \in
S^c$, for all $i=1,\ldots,n$,
\begin{equation}
\lambda_n \geq \mathsf{range}_k \Big| \frac{32}{n}\hat{r}^\tran \mathbf{1}_{(i':n)_k} \Big| ,
\end{equation}
$\max_{i'} \frac{X_{k\pi_k(i'+1)} - X_{k \pi_k(i')}}{\mathsf{range}_k} \leq \frac{1}{16}$, $\mathsf{range}_k \geq 1$, and both the AC solution $\hat{f}_k$ from optimization~\eqref{opt:alternate_opt} and the DC solution $\hat{g}_k$ from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}

>>>>

In our analysis, we assume that an upper bound $B$ to $\| \hat{f}_k \|_\infty$ 
is imposed in the optimization procedure, where $B$ is chosen to also upper bound $\| f^*_k \|_\infty$ (same $B$ as in Assumption A3 in Section~\ref{sec:probabilistic_setting}). This $B$-boundedness constraint is added so that we may use the convex function bracketing results from~\cite{kim2014global} to establish uniform convergence between the empirical risk and the population risk. We emphasize that this constraint is not needed in practice and we do not use it for any of our simulations.

@@

Theorem~\ref{thm:deterministic} allows us to separately analyze the false negative
rates and false positive rates. To control false positives,
Theorem~\ref{thm:false_positive} verifies that the conditions in Theorem~\ref{thm:deterministic} hold in a stochastic setting. To control
false negatives, Theorem~\ref{thm:false_negative} analyzes the restricted regression with only $|S|$ variables.

The proof of Theorem~\ref{thm:deterministic} analyses the KKT
conditions of optimization~\eqref{opt:alternate_opt}.  This parallels
the now standard \emph{primal-dual witness}
technique~\citep{wainwright2009sharp}. The conditions in Theorem~\ref{thm:deterministic} is our analogue to the \emph{mutual incoherence} condition. Our conditions are much more strict however because the estimation is
nonparametric---even the low dimensional restricted regression has
$s(n-1)$ variables.

The details of the proof are given in
Section~\ref{sec:deterministic_proof} of the Appendix.

\subsection{Probabilistic Setting}
\label{sec:probabilistic_setting}

@@

\begin{theorem}[Controlling false positives]
\label{thm:false_positive}
Suppose assumptions A1-A5 hold. Define $\tilde{\sigma} \equiv \max(\sigma, B)$ and define $\mathsf{range}_k = X_{\pi_k(n)k} - X_{\pi_k(1)k}$. Suppose $p \leq O\big( \exp(c n) \big)$ and $n \geq C$ for some positive constants $0 < c <1$ and $C$.  Suppose also
\begin{equation}
\lambda_n \geq 768 s \tilde{\sigma}  \sqrt{ \frac{\log^2 np}{n}}.
\end{equation}  
Then with probability at least $ 1 - \frac{24}{n}$, for all $k \in
S^c$, for all $i=1,\ldots,n$,
\begin{equation}
\lambda_n \geq \mathsf{range}_k \Big| \frac{32}{n}\hat{r}^\tran \mathbf{1}_{(i':n)_k} \Big| ,
\end{equation}
$\max_{i'} \frac{X_{\pi_k(i'+1)k} - X_{\pi_k(i')k}}{\mathsf{range}_k} \leq \frac{1}{16}$, $\mathsf{range}_k \geq 1$, and both the AC solution $\hat{f}_k$ from optimization~\eqref{opt:alternate_opt} and the DC solution $\hat{g}_k$ from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}
